---
title: "SPA Running Completion Report"
author: "SPA Team"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: kate
    df_print: kable

subtitle: Tracking completion and engagement
mainfont: Univers
---
<style type="text/css">
   .main-container {max-width: 70%;}
   .row {display: flex;}
   .column {flex: 50%;}
</style>


```{r setup, include=FALSE}
# include=FALSE means it won't include any code blocks in the knit
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
rm(list=ls()) 
```

```{r}
# libraries
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(ggpattern)
library(patchwork)
library(tidyr)
library(gridExtra)
library(scales)  # For percent labels on the y-axis
```

```{r, fig.width=12, fig.height=8}
# show options for colours from Brewer's Palette
display.brewer.all()
```


```{r}
# set colours
display.brewer.pal(9, "Blues")
# display.brewer.pal(9, "Set2")

```





```{r}
# Load file with completion stats (produced in Python)
engagement_stats_o <- read.csv("overall_engagement_stats_by_date.csv", 
                               row.names = 1, check.names = FALSE)


engagement_stats_s <- read.csv("engagement_stats_by_school.csv", 
                               row.names = 1, check.names = FALSE)

total_completion_stats = read.csv("overall_completion_rates.csv")
colnames(total_completion_stats)[1] <- "Status"
colnames(total_completion_stats)[2] <- "Candidates"
```

<style type="text/css">
   .main-container {max-width: 100%;}
   .row {display: flex;}
   .column {flex: 50%;}


## How many candidates have completed the exam in total?



```{r}

df <- total_completion_stats %>%
  filter(Status != "Percentage Completed") %>%   # Remove the percentage row
  mutate(Candidates = as.numeric(Candidates))    # Convert to numeric

# Create the pie chart using the filtered dataframe
pie_chart <- ggplot(df, aes(x = "", y = Candidates, fill = Status)) +
  geom_bar(stat = "identity", width = 1, color = "white") + 
  coord_polar("y", start = 0) + 
  scale_fill_manual(values = c("#2E8B57", "#A1D99B")) +  # Gradient-like colors
  theme_void() + 
  labs(title = "Completion Status") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14)) + 
  geom_text(aes(label = paste0(round((Candidates / sum(Candidates)) * 100, 1), "%")),
            position = position_stack(vjust = 0.5), 
            size = 6, fontface = "bold", color = "white")


```
</style>
<div class = "row">
<div class = "column">
### Total completion statistics
```{r echo=FALSE}
```


```{r echo=FALSE}
pie_chart
```
</div>

<div class = "column">
```{r echo=FALSE}
total_completion_stats
```
</div>
</div>


```{r, fig.width=15, fig.height=6, fig.align='center'}
# Table of engagement stats overall
engagement_stats_o

# Transpose the data so dates are in one column
engagement_long <- as.data.frame(t(engagement_stats_o))
engagement_long$Date <- rownames(engagement_long)

# Convert to long format
engagement_long <- pivot_longer(engagement_long, 
                                cols = -Date, 
                                names_to = "Category", 
                                values_to = "Percentage")

# Convert Date to proper format
engagement_long$Date <- as.Date(engagement_long$Date)

# Define a theme with larger text
big_text_theme <- theme(
  plot.title = element_text(size = 45, face = "bold"),  # Bigger title
  axis.title = element_text(size = 25),                # Bigger axis labels
  axis.text = element_text(size = 15),                 # Bigger axis text
  legend.title = element_text(size = 25, face = "bold"), # Bigger legend title
  legend.text = element_text(size = 20)                # Bigger legend text
)

# Plot three separate line graphs (one per row)
ggplot(engagement_long, aes(x = Date, y = Percentage, color = Category)) +
    geom_line(size = 1) +
    facet_grid(Category ~ ., scales = "free_y") +  # Forces each graph to be on a separate row
    labs(title = "Engagement Statistics Over Time", 
         x = "Date", y = "Percentage") +
    theme_minimal() + big_text_theme

```


<hr>




## Engagement rates by high school
<hr>

```{r}
# Remove any leading "X" from column names and convert them to Date objects
colnames(engagement_stats_o) <- as.Date(gsub("^X", "", colnames(engagement_stats_o)), 
                                      format = "%Y-%m-%d")

# Save each row (metric) as a variable
test_time_low_by_date    <- as.numeric(engagement_stats_o["Low total test time percentage", ])
writing_time_low_by_date <- as.numeric(engagement_stats_o["Low writing time percentage", ])
writing_words_low_by_date<- as.numeric(engagement_stats_o["Low word count percentage", ])

# Create data frames for plotting.
# Each data frame has a 'Date' column (from the column names) and a 'Percentage' column (the row values)
df_test_time <- data.frame(Date = as.Date(colnames(engagement_stats_o)),
                           Percentage = test_time_low_by_date)
df_writing_time <- data.frame(Date = as.Date(colnames(engagement_stats_o)),
                              Percentage = writing_time_low_by_date)
df_writing_words <- data.frame(Date = as.Date(colnames(engagement_stats_o)),
                               Percentage = writing_words_low_by_date)

p1 <- ggplot(df_test_time, aes(x = Date, y = Percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Completed test in < 20% of time given", x = "Date", y = "Percentage") +
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal()

# Create bar chart for Low Writing Time Percentage
p2 <- ggplot(df_writing_time, aes(x = Date, y = Percentage)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  labs(title = "Completed writing in < 10% of time given", x = "Date", y = "Percentage") +
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal()

# Create bar chart for Low Word Count Percentage
p3 <- ggplot(df_writing_words, aes(x = Date, y = Percentage)) +
  geom_bar(stat = "identity", fill = "darkred") +
  labs(title = "Wrote < 20 words", x = "Date", y = "Percentage") +
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal()

# Print the plots
p1
p2
p3

```


## Student engagement by school

The following are strong evidence that candidates are not engaging fully in the test:

- Candidates who used less than **20%** of the total allocated test time are likely not engaging fully in the assessment.
- Candidates who used less than **10%** of the allocated writing time.
- Candidates who wrote less than **20 words** on the writing section.

 
```{r, fig.width=6, fig.height=6, fig.align='center', fig.show='hold'}
engagement_stats_s


```

